---
title: "PenguinFlipperLengthAnova"
author: "Madeline Gorchels"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##1.Loading the packages and data 
```{r, include = FALSE}
library(tidyverse)
library(effsize)
library(car)
library(onewaytests)

penguins = read_csv("penguins.csv")

```

##2. Exploritory Data 

```{r}

```

##3. Is there a significant difference in mean flipper lengths for female Chinstrap, Gentoo, and Adelie Penguins near Palmer Archipelago. 

Maybe I can use a one way ANOVA. 
1.Assuming we have independent observations
2.Equal variances
3.Assumption of normality

```{r}
#Use the table function to figure out the counts for each level
counts = with(penguins, table(species, sex))

#View(counts)

#should always look to see if the mean is a good parameter 
penguins2 = penguins %>% 
  mutate(
    common_name = 
      case_when(
        species =="Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap", 
        species == "Gentoo penguin (Pygoscelis papua)" ~ "Genoo",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")

#View(penguins2)
```

```{r}
hists = ggplot(penguins2, aes(x = flipper_length))+
  geom_histogram(aes(fill = common_name))+
  scale_fill_grey()+
  facet_wrap(~ common_name, scale = "free")+
  theme_classic()+
  scale_y_continuous(expand = c(0,0))+ 
  scale_x_continuous(expand = c(0,0))

hists

qq = ggplot(penguins2, aes(sample = flipper_length))+
  geom_qq(aes(color = common_name))+
  scale_fill_grey()+
  facet_wrap(~ common_name, scale = "free")+
  theme_classic()

qq

#Based on qq plots and histograms observations for each group seem normally distributed. Also n > 30 so the central limit theorem applies, meaning the means will be normally distributed. 
```

##4. Evaulate variances and preform a Levine's test for equal variances

```{r}
# If the largest sample variance is 4x greater than the smallest variance then thse are "close enough" to consider variances equal

variances = penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = round(var(flipper_length), 2)
  )

variances #Since the largest is only twice the smallest we can assume the variances are the same 

#more than 2 samples: use Levene's test (from the car package)
#levenTest function

var_test = leveneTest(penguins2$flipper_length ~ penguins2$common_name)

var_test

#H0: No differences in variances 
#HA: Variances are not eqaul 

#There is no significant difference in variances across species.  
```

##5. One-way ANOVA for penguin flipper length by species (common_name)

```{r}

# H0: means accross all groups are eqaul 
# HA: at least two group means differ significantly 

penguin_aov = aov(penguins2$flipper_length~penguins2$common_name)

summary(penguin_aov)

#post-hoc tesing by Tukey's HSD
# H0 : no significant difference in pairwise means 
# HA : is a significant difference in pairwise means 
penguin_ph = TukeyHSD(penguin_aov)

penguin_ph

#aov.test function does allow for inline referencing and has lots of post hoc test options 

```

Mean flipper length of female penguins differed signfificantly for all species studied (one-way anova, F(df), p < 0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha$ = 0.05)
